---
title: "GLMM - Pass Analysis"
output: html_notebook
---

```{r, message=F, warning=F}
library(dplyr)
library(ggplot2)
library(glmmTMB)
library(lubridate)
library(nadiv)
```

```{r}
load = read.csv(paste0("C:/Users/Ong Jia Yi/Desktop/STUDY/Summer 2021/",
                      "Rotman MMA Summer 2021 Datathon/",
                      "NWHL_1.csv"),
                stringsAsFactors = FALSE)
```



# PASS ANALYSIS

## Helper Functions

```{r}
compute_distance = function(x0, y0, x1, y1){
  
  distance = sqrt((x1 - x0)^2 + (y1 - y0)^2)
  
  return(distance)
}
```

```{r}
compute_angle_pass = function(x0, y0, x1, y1, team){
  # Returns angle between direction of pass and 
  # direction of offense of the player's team.
  # Outputs angle in degrees.
  # Define passing on the spot to have zero angle.
  
  pass_dir = c(x1 - x0, y1 - y0)
  pass_mag = sqrt(sum(pass_dir^2))
  
  if (team == "home"){
    off_dir = c(1, 0)
  }
  if (team == "away"){
    off_dir = c(-1, 0)
  }
  
  angle = ifelse(pass_mag != 0, 
                 acos(sum(off_dir * pass_dir) / pass_mag),
                 0)
  
  return(angle/(2*pi)*360)
}
```


## Data Preprocessing

```{r}
data_pass = load %>%
  filter(Event %in% c("Incomplete Play", "Play")) %>% 
  
  # data formatting
  mutate(Period = factor(Period, levels = c(1, 2, 3, 4)),
         Clock = ms(Clock),
         time_played = hms(substr(Cumulative_Time_Play, 8, 1000)),
         direct_pass = ifelse(Detail.1 == "Direct", 1, 0)) %>%
  
  # create angle and distance variables
  mutate(status = factor(ifelse(Team == Home.Team, "home", "away"))) %>%
  rowwise() %>%
  mutate(pass_angle = compute_angle_pass(X.Coordinate, Y.Coordinate,
                                         X.Coordinate.2, Y.Coordinate.2,
                                         status),
         pass_dist = compute_distance(X.Coordinate, Y.Coordinate,
                                      X.Coordinate.2, Y.Coordinate.2)) %>%
  
  # create pair variable
  mutate(pairs = ifelse(Player < Player.2,
                        paste0(Player, " - ", Player.2),
                        paste0(Player.2, " - ", Player))) %>%
  
  # create own/opponent team score
  mutate(own_score = ifelse(Team == Home.Team, Home.Team.Goals, Away.Team.Goals),
         opponent_score = ifelse(Team == Home.Team, Away.Team.Goals, Home.Team.Goals)) %>%
  
  # create strength status
  mutate(strength_status = case_when(
    Home.Team.Skaters - Away.Team.Skaters == 0 ~ "Even_Strength",
    Home.Team.Skaters - Away.Team.Skaters == 1 ~ ifelse(Team == Home.Team, "Power_Play_1", "Penalty_Kill_1"),
    Home.Team.Skaters - Away.Team.Skaters == 2 ~ ifelse(Team == Home.Team, "Power_Play_2", "Penalty_Kill_2"),
    Home.Team.Skaters - Away.Team.Skaters == -1 ~ ifelse(Team != Home.Team, "Power_Play_1", "Penalty_Kill_1"),
    Home.Team.Skaters - Away.Team.Skaters == -2 ~ ifelse(Team != Home.Team, "Power_Play_2", "Penalty_Kill_2"))) %>%
  
  # create relative X,Y variables
  mutate(X = ifelse(Team == Home.Team, X.Coordinate, abs(200 - X.Coordinate)),
         X.2 = ifelse(Team == Home.Team, X.Coordinate.2, abs(200 - X.Coordinate.2))) %>%
  rename(Y = Y.Coordinate, Y.2 = Y.Coordinate.2) %>%
  
  # create target variable
  mutate(target = ifelse(Event == "Play", 1, 0)) %>%
  
  select(Team, pairs, Period, Clock, Y, Y.2, X, X.2,
         time_played, direct_pass, status, pass_angle,
         pass_dist, own_score, opponent_score, strength_status,
         target)
```

```{r}
data_pass
```


## Visualization

Verifying angle and distance computation
```{r}
plot_data = data_pass %>%
  filter(target == 1,
         direct_pass == 1,
         status == "away",
         Team == unique(data_pass$Team)[1])

ggplot(plot_data[1:20,]) +
  geom_segment(aes(x = X, y = Y, xend = X.2, yend = Y.2),
               arrow = arrow()) +
  geom_text(aes(x = (X + X.2)/2, y = (Y + Y.2)/2,
                label = paste(as.character(round(pass_angle, 2)),
                              ", ",
                              as.character(round(pass_dist, 2)))),
            size = 3)
```


## Model Fitting

```{r}
t0 = Sys.time()
model1_full = glmmTMB::glmmTMB(target ~ Period + Clock + time_played + status
                               + direct_pass + pass_angle + pass_dist 
                               + own_score + opponent_score + strength_status
                               + Y + Y.2 + X + X.2
                               + (1 | Team/pairs), 
                               family = binomial(link = "logit"), 
                               data = data_pass)

model1_noTeam = glmmTMB::glmmTMB(target ~ Period + Clock + time_played + status
                                 + direct_pass + pass_angle + pass_dist 
                                 + own_score + opponent_score + strength_status
                                 + Y + Y.2 + X + X.2
                                 + (1 | pairs), 
                                 family = binomial(link = "logit"), 
                                 data = data_pass)

model1_none = glmmTMB::glmmTMB(target ~ Period + Clock + time_played + status
                               + direct_pass + pass_angle + pass_dist 
                               + own_score + opponent_score + strength_status
                               + Y + Y.2 + X + X.2, 
                               family = binomial(link = "logit"), 
                               data = data_pass)
t1 = Sys.time()
t1 - t0
```


## Model Selection

Likelihood Ratio Test:
Null hypothesis that the smaller models provides as good a fit for the data as the larger model
REJECTED.
```{r}
nadiv::LRTest(logLik(model1_full), logLik(model1_noTeam), 
              df = 1, boundaryCorrection = TRUE)$Pval
nadiv::LRTest(logLik(model1_full), logLik(model1_none), 
              df = 2, boundaryCorrection = TRUE)$Pval
```

```{r}
anova = anova(model1_full, model1_noTeam, model1_none)
anova
```


## Final Model

```{r}
summary(model1_full)
```


## Model Evaluation

```{r, message=F, warning=F}
library(caret)
library(pROC)
library(plotROC)
```

```{r}
pred = factor(ifelse(fitted(model1_full) >= 0.5, 1, 0), levels = c(1, 0))
y = factor(data_pass$target, levels = c(1, 0))
```

```{r}
confusionMatrix(pred, y)
```

```{r, message=F}
invisible(plot(roc(y, fitted(model1_full)), print.thres = c(.1, .5),
               col = "red", print.auc = T))
```


## Interpreting Model

Displaying odds-ratios
```{r}
confint(model1_full) %>% exp()
```

```{r}
ranef = glmmTMB::ranef(model1_full)
ranef$cond$`pairs:Team` %>%
  arrange(desc(`(Intercept)`))
```

Raw success rate for comparison
```{r}
data_pass %>%
  group_by(pairs) %>%
  summarise(counts = n(), rate = sum(target)/counts) %>%
  filter(counts > 20) %>%
  arrange(desc(counts), desc(rate))
```

```{r}
ranef$cond$Team %>%
  arrange(desc(`(Intercept)`))
```


