---
title: "GLMM - Shot Analysis"
output: html_notebook
---

```{r, message=F, warning=F}
library(dplyr)
library(ggplot2)
library(glmmTMB)
library(lubridate)
library(nadiv)
```

```{r}
load = read.csv(paste0("C:/Users/Ong Jia Yi/Desktop/STUDY/Summer 2021/",
                      "Rotman MMA Summer 2021 Datathon/",
                      "NWHL_1.csv"),
                stringsAsFactors = FALSE)
```



# SHOT ANALYSIS

## Helper Functions

```{r}
compute_distance_shot = function(x, y){
  
  distance = sqrt((200 - x)^2 + (85/2 - y)^2)
  
  return(distance)
}
```

```{r}
compute_angle_shot = function(x, y){
  # angle from direction facing the net
  
  shot_vector = c(x - 200, y - 85/2)
  shot_mag = sqrt(sum(shot_vector^2))
  
  radian = ifelse(shot_mag != 0, 
                  acos(sum(c(1, 0) * shot_vector) / shot_mag),
                  0)
  angle = radian/(2*pi)*360
  
  return(ifelse(angle > 90, 180-angle, angle))
}
```


## Data Preprocessing

```{r}
data_shot = load %>%
  filter(Event %in% c("Shot", "Goal")) %>% 
  
  # data formatting
  mutate(Period = factor(Period, levels = c(1, 2, 3, 4)),
         Clock = ms(Clock),
         time_played = hms(substr(Cumulative_Time_Play, 8, 1000)),
         Detail.1 = as.factor(Detail.1),
         Detail.3 = factor(Detail.3, levels = c('f', 't')),
         Detail.4 = factor(Detail.4, levels = c('f', 't'))) %>%
  
  # create angle and distance variables
  mutate(status = factor(ifelse(Team == Home.Team, "home", "away"))) %>%
  rowwise() %>%
  mutate(shot_angle = compute_angle_shot(X.Coordinate, Y.Coordinate),
         shot_dist = compute_distance_shot(X.Coordinate, Y.Coordinate)) %>%
  
  # create own/opponent team score
  mutate(own_score = ifelse(Team == Home.Team, Home.Team.Goals, Away.Team.Goals),
         opponent_score = ifelse(Team == Home.Team, Away.Team.Goals, Home.Team.Goals)) %>%
  
  # create strength status
  mutate(strength_status = case_when(
    Home.Team.Skaters - Away.Team.Skaters == 0 ~ "Even_Strength",
    Home.Team.Skaters - Away.Team.Skaters == 1 ~ ifelse(Team == Home.Team, "Power_Play_1", "Penalty_Kill_1"),
    Home.Team.Skaters - Away.Team.Skaters == 2 ~ ifelse(Team == Home.Team, "Power_Play_2", "Penalty_Kill_2"),
    Home.Team.Skaters - Away.Team.Skaters == -1 ~ ifelse(Team != Home.Team, "Power_Play_1", "Penalty_Kill_1"),
    Home.Team.Skaters - Away.Team.Skaters == -2 ~ ifelse(Team != Home.Team, "Power_Play_2", "Penalty_Kill_2"))) %>%
  
  rename(X = X.Coordinate, Y = Y.Coordinate) %>%
  
  # create target variable
  mutate(target = ifelse(Event == "Goal", 1, 0)) %>%
  
  select(Team, Player, Period, Clock, Y, X,
         time_played, status, shot_angle, shot_dist, own_score, 
         Detail.1, Detail.3, Detail.4, opponent_score, strength_status,
         target)
```

```{r}
write.csv(data_shot, "data_shot.csv", row.names=FALSE)
```



## Visualization

Verifying angle and distance computation
```{r}
plot_data = data_shot %>%
  filter(target == 1,
         Team == unique(data_shot$Team)[1])

ggplot(plot_data[1:10,]) +
  geom_segment(aes(x = X, y = Y, xend = 200, yend = 85/2),
               arrow = arrow()) +
  geom_text(aes(x = X, y = Y,
                label = paste(as.character(round(shot_angle, 2)),
                              ", ",
                              as.character(round(shot_dist, 2)))),
            size = 3)
```


## Model Fitting

```{r}
t0 = Sys.time()
model1_full = glmmTMB::glmmTMB(target ~ Clock + time_played + status
                               + shot_angle + shot_dist 
                               + own_score + opponent_score + strength_status
                               + Detail.4
                               + (1 | Team/Player), 
                               family = binomial(link = "logit"), 
                               data = data_shot)

model1_noTeam = glmmTMB::glmmTMB(target ~ Clock + time_played + status
                                 + shot_angle + shot_dist 
                                 + own_score + opponent_score + strength_status
                                 + Detail.4
                                 + (1 | Player), 
                                 family = binomial(link = "logit"), 
                                 data = data_shot)

model1_none= glmmTMB::glmmTMB(target ~ Clock + time_played + status
                              + shot_angle + shot_dist 
                              + own_score + opponent_score + strength_status
                              + Detail.4, 
                              family = binomial(link = "logit"), 
                              data = data_shot)
t1 = Sys.time()
t1 - t0
```


## Model Selection

Likelihood Ratio Test:
Null hypothesis that the smaller models provides as good a fit for the data as the larger model
REJECTED.
```{r}
nadiv::LRTest(logLik(model1_full), logLik(model1_noTeam), 
              df = 1, boundaryCorrection = TRUE)$Pval
nadiv::LRTest(logLik(model1_full), logLik(model1_none), 
              df = 2, boundaryCorrection = TRUE)$Pval
```

```{r}
anova = anova(model1_full, model1_noTeam, model1_none)
anova
```


## Final Model

```{r}
summary(model1_full)
```


## Model Evaluation

```{r, message=F, warning=F}
library(caret)
library(pROC)
library(plotROC)
```

```{r}
pred = factor(ifelse(fitted(model1_full) >= 0.5, 1, 0), levels = c(1, 0))
y = factor(data_shot$target, levels = c(1, 0))
```

```{r}
confusionMatrix(pred, y)
```

```{r, message=F}
invisible(plot(roc(y, fitted(model1_full)), print.thres = c(.1, .5),
               col = "red", print.auc = T))
```


## Interpreting Model

Displaying odds-ratios
```{r}
confint(model1_full) %>% exp()
```

```{r}
ranef = glmmTMB::ranef(model1_full)
ranef$cond$`Player:Team` %>%
  arrange(desc(`(Intercept)`))
```

Raw success rate for comparison
```{r, message=F}
data_shot %>%
  group_by(Player) %>%
  summarise(counts = n(), rate = sum(target)/counts) %>%
  filter(counts > 20) %>%
  arrange(desc(counts), desc(rate))
```

```{r}
ranef$cond$Team %>%
  arrange(desc(`(Intercept)`))
```


